Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.0

### 2019-03-02 manuscript.Rmd
cover letter, text updates

---
title: 'Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.00'
author:
  - Vinay Swamy:
      institute: ogvfb
  - David McGaughey:
      institute:
        - ogvfb
      correspondence: "yes"
      email: mcgaugheyd@mail.nih.gov
institute:
  - ogvfb: Ophthalmic Genetics and Visual Function Branch, National Eye Institute, National Institutes of Health
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: eyeIntegration_v1_app_manuscript.bib
csl: investigative-ophthalmology-and-visual-science.csl
abstract: "PURPOSE: To provide the eye community with an easily accessible and reliable RNA-seq based transcriptome expression database of healthy human eye tissues and a matching reactive web application to query gene expression in eye and body tissues. METHODS: We downloaded the raw sequnce data for 892 RNA-seq based samples across 44 tissues in the GTEx project as a non-eye reference set. We then queried several public repositories to find all healthy, non-perturbed, human eye tissue RNA-seq samples. The 1311 total samples we found are sent into a Snakemake-based reproducible pipeline we wrote. The pipeline automatically quantifies all known transcripts and genes, removes samples with poor sequence quality and mislabels, normalizes expression values across each tissue, performs 105 differential expression tests, calculates GO term enrichment, and outputs all as a single SQLite database file: the Eye in a Disk (EiaD) dataset. RESULTS: As 40+ GB database is not a useable format for most, we rewrote the web application  eyeIntegration (version 1.02 as of Februay 24 2019, https://eyeIntegration.nei.nih.gov) to display these data, focusing on providing users quick visualizations by database version, gene/transcript, and 60 tissues. As a demonstration of the value of this unified pan-eye dataset, we show that fetal and organoid retina tissues are highly similar at a pan-transcriptome level but contain several gene set differences that suggest targets for organoid process improvements. CONCLUSION: The eyeIntegration v1.02 web app serves the pan-eye and body transcriptome dataset, EiaD. This provides the eye community a powerful and quick mean to testing hypothesis relating to human gene and transcript expression across 44 body and 16 eye tissues."
keywords: "RNA-seq, lens, cornea, retina, RPE, Snakemake, web, app <- Not used in Word as far as I can tell"
---

>>>>

---
title: 'Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.0'
author:
  - Vinay Swamy:
      institute: ogvfb
  - David McGaughey:
      institute:
        - ogvfb
      correspondence: "yes"
      email: mcgaugheyd@mail.nih.gov
institute:
  - ogvfb: Ophthalmic Genetics and Visual Function Branch, National Eye Institute, National Institutes of Health
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: eyeIntegration_v1_app_manuscript.bib
csl: investigative-ophthalmology-and-visual-science.csl
abstract: "PURPOSE: To provide the eye community with an easily accessible and reliable RNA-seq based transcriptome expression database of healthy human eye tissues and a matching reactive web application to query gene expression in eye and body tissues. METHODS: We downloaded the raw sequnce data for 892 RNA-seq based samples across 44 tissues in the GTEx project as a non-eye reference set. We then queried several public repositories to find all healthy, non-perturbed, human eye tissue RNA-seq samples. The 1311 total samples we found are sent into a Snakemake-based reproducible pipeline we wrote. The pipeline automatically quantifies all known transcripts and genes, removes samples with poor sequence quality and mislabels, normalizes expression values across each tissue, performs 105 differential expression tests, calculates GO term enrichment, and outputs all as a single SQLite database file: the Eye in a Disk (EiaD) dataset. RESULTS: As 40+ GB database is not a useable format for most, we rewrote the web application  eyeIntegration (https://eyeIntegration.nei.nih.gov) to display these data, focusing on providing quick visualizations by database version, gene/transcript, and 60 tissues. As a demonstration of the value of this unified pan-eye dataset, we show that fetal and organoid retina tissues are highly similar at a pan-transcriptome level but contain several gene set differences that suggest organoid process improvements. CONCLUSION: The eyeIntegration v1.0 web app serves the pan-eye and body transcriptome dataset, EiaD. This provides the eye community a powerful and quick mean to testing hypothesis relating to human gene and transcript expression across 44 body and 16 eye tissues."
keywords: "RNA-seq, lens, cornea, retina, RPE, Snakemake, web, app <- Not used in Word as far as I can tell"
---
